---
title: "Lab Notebook"
author: "EShelton"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# First Steps

My directory is found at: /projectnb/vervpop/eshelton. (Input: "cd /projectnb/vervpop" "cd eshelton")

The location of *PPARGC1A*, or *PGC-1α*, in the Vervet genome is **27: 26,465,499-26,564,582**. With a buffer zone of 10k base pairs, the location is **27: 26,455,499-26,574,582**.

## Getting the gene region:

```{r}`r''`
wget https://vervetmonkeygmi.blob.core.windows.net/publicationdata/163_201701_UG_vervet_monkey_SNPs_all_chrom_beagle_shapeit.vcf.gz

tabix -p vcf 163_201701_UG_vervet_monkey_SNPs_all_chrom_beagle_shapeit.vcf.gz.1

tabix -h 163_201701_UG_vervet_monkey_SNPs_all_chrom_beagle_shapeit.vcf.gz.1 CAE27:26455499-26574582 > vervPGC1a.full.vcf

less vervPGC1a.full.vcf
```
```
##fileformat=VCFv4.1
##FILTER=<ID=PASS,Description="All filters passed">
##fileDate=14022017_09h11m56s
##source=SHAPEIT2.v837
##log_file=shapeit_14022017_09h11m56s_34f53654-f41e-4bf1-89fe-892d6199343a.log
##FORMAT=<ID=GT,Number=1,Type=String,Description="Phased Genotype">
##contig=<ID=CAE1>
##contig=<ID=CAE2>
##contig=<ID=CAE3>
##contig=<ID=CAE4>
##contig=<ID=CAE5>
##contig=<ID=CAE6>
##contig=<ID=CAE7>
##contig=<ID=CAE8>
##contig=<ID=CAE9>
##contig=<ID=CAE10>
##contig=<ID=CAE11>
##contig=<ID=CAE12>
##contig=<ID=CAE13>
##contig=<ID=CAE14>
##contig=<ID=CAE15>
##contig=<ID=CAE16>
##contig=<ID=CAE17>
##contig=<ID=CAE18>
vervPGC1a.full.vcf
```
Hit 'q' to close the less command view.
```{r}`r''`
wc -l vervPGC1a.full.vcf
```
```
2340 vervPGC1a.full.vcf
```
```{r}`r''`
grep -c "^##" vervPGC1a.full.vcf
```
```
38  
```

Note: At this point vervPGC1a.full.vcf was renamed to vervpcg1a.full.vcf.

Questions:  
If the download of a file from the SCC is interrupted accidentally, how do I unpause the download without starting it all over again?     
When I used the 'tabix -p' command, when I hit return it just went down a line; after a couple minutes it goes. How do I make it run sooner?  

## Functional Regions

Number: Start coordinate-End coordinate (length)

**Exon 1: 26,465,499-26,465,783 (285)**  
  *Intron 1-2: 26465784-26470781 (4998)*   
**Exon 2: 26470782-26470961 (180)**   
  *Intron 2-3: 26470962-26523412 (52451)*   
**Exon 3: 26523413-26523607 (195)**   
  *Intron 3-4: 26523608-26525582 (1975)*   
**Exon 4: 26525583-26525705 (123)**   
  *Intron 4-5: 26525706-26526565 (860)*   
**Exon 5: 26526566-26526770 (205)**   
  *Intron 5-6: 26526771-26530747 (3977)*   
**Exon 6: 26530748-26530793 (46)**   
  *Intron 6-7: 26530794-26530902 (109)*   
**Exon 7: 26530903-26530976 (74)**   
  *Intron 7-8: 26530977-26540652 (9676)*   
**Exon 8: 26540653-26541568 (916)**    
  *Intron 8-9: 26541569-26542129 (561)*   
**Exon 9: 26542130-26542234 (105)**   
  *Intron 9-10: 26542235-26542387 (153)*   
**Exon 10: 26542388-26542508 (121)**   
  *Intron 10-11: 26542509-26552948 (10440)*   
**Exon 11: 25552949-26553070 (122)**   
  *Intron 11-12: 26553071-26553415 (345)*   
**Exon 12: 26553416-26553567 (152)**   
  *Intron 12-13: 26553568-26559651 (6084)*   
**Exon 13: 26559652-26564582 (4931)**

## Hardy-Weinburg Equilibrium

I ran into trouble with the 'vervet.population.panel' file required for this section. It wasn't in the vervpop directory, so Christian Gagnon sent it to me and I downloaded it onto my computer. I then had issues with uploading the file to the SCC. The SCC cheat sheet pdf used the command 'scp *filename* *username*@scc1.bu.edu:~', but when I ran that command the terminal said that no such file or directory existed. I tried it in my home directory, the vervpop directory, and my eshelton directory. I was able to upload the file using SCC On Demand.

I created the file 'pygpgc1a.vcf' of the variants within the South African *pygerythrus* sample using the given code, deleting the VSAI3005 sample. 

```{r}`r''`
grep pygerythrus vervet.population.panel | cut -f1 > pyg.samples.list
gedit pyg.samples.list
module load vcftools
vcftools --vcf vervUCP1.full.vcf --keep pyg.samples.list --recode --out pygUCP1
mv pygUCP1.recode.vcf pygUCP1.vcf
```

I created the file 'sepgc1a.vcf' of the variants within the Southern Expansion using the given code. I left the VCSAI3005 sample in. The instructions also mentioned changing the taxonomic designation of two of the samples, which I was unsure how to do.

# SNPs

I opened an SCC OnDemand R server, and used the R code given in the pipeline to analyze the 'pygpgc1a.hwe' file. It returned 102 SNPs. The 'sepgc1a.hwe' file returned 276 SNPs. The details of these have been copied to text files. 

There was a complication and Christian Gagnon sent me the correct version of the 'vervet.population.panel' file. I redid the creation of the Pygerythrus and Southern Expansion files, 'pygpgc1a.hwe' and 'sepgc1a.hwe', making sure to edit 'vervet.population.panel' file in between to make the changes necessary to the VBOA and VSAJ2008 samples.
